Identification of protoneutron star g-modes in gravitational-wave data

### 2020-08-31 methods.tex
minor fixes

In this section, we outline a strategy for estimating the time evolution of the
ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the PNS and its squared radius
(in units of solar mass and km) from the observation of the $\mbox{}^2g_2$
oscillation mode in the gravitational wave detector data.
An integral part of this strategy is the universal relations that relate the
characteristic frequency of the PNS oscillation $f$, $g$ and $p$ modes with the mass
and the radius of the PNS, the shock radius and the total mass inside the shock as
demonstrated in \cite{Torres:2019}.

@@

To develop the method we considered the gravitational wave signal
{\tt s20-gw-10kpc} described in Section \ref{sec:simulations}, originally
sampled at 16384 Hz but resampled at 4096 Hz.
A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
autoregressive estimates of the local spectra for successive time intervals of 
\textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.
The dominant emission mode corresponds to the PNS oscillation $\mbox{}^2 g_2$-mode. We have
developed a time-frequency method to track the ridge $m(t)$ in the spectrogram,
taking into account that it is monotonically increasing as time goes,
a property of the $\mbox{}^2 g_2$-mode.
Starting from either the left- or right-most column of the time-frequency matrix
we identify and trace the sequence of amplitude peaks within a certain frequency
band given the monotonicity constraint. Appendix \ref{app:gmode} is providing more
details on the reconstruction of the $g$ mode ridge.

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{plots/spectrogram}
 \caption{Spectrogram of the gravitational wave signal {\tt s20-gw-10kpc} sampled at \unit[4096]{Hz}.
   The spectrogram is obtained using data streach of 200 samples overlapping at 90\%
   with each other.} \label{fig:spectrogram}
\end{figure}

We collect the instantaneous frequency $f(t_i)$ corresponding to the ridge $m(t_i)$ for
the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
for values in between the $t_i$. We then use equation (\ref{eq:universal}) to obtain
estimates of the time evolution of the ratio together with 95\% confidence intervals.
An exemple is given in Figure \ref{fig:ratio} where the red points are the point estimates and
the grey bands represent 95\% confidence bands. Ratio values
computed using the mass and radius values obtained from the simulation code are shown in black.

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/ratio}
 \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ as function of time extracted from the $\mbox{}^2 g_2$-mode of the {\tt s20-gw-10kpc} signal (red points and the 95\% confidence belt in grey) compared to the ratio value derived from the PNS mass and radius given by the simulation code (black points).} \label{fig:ratio}
\end{figure}

>>>>

In this section, we outline a strategy for estimating the time evolution of the
ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the PNS and its squared radius
(in units of solar mass and km) from the observation of the $\mbox{}^2g_2$
oscillation mode in the gravitational wave detector data.
An integral part of this strategy is the universal relations that relate the
characteristic frequency of the PNS oscillation $f$, $g$ and $p$ modes with the mass
and the radius of the PNS, the shock radius and the total mass inside the shock as
demonstrated in \cite{Torres:2019b}.

@@

To develop the method we considered the gravitational wave signal
{\texttt s20S} described in Section \ref{sec:simulations}, originally
sampled at 16384 Hz but resampled at 4096 Hz.
A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
autoregressive estimates of the local spectra for successive time intervals of 
\textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.
The dominant emission mode corresponds to the PNS oscillation $\mbox{}^2 g_2$-mode. We have
developed a time-frequency method to track the ridge $m(t)$ in the spectrogram,
taking into account that it is monotonically increasing as time goes,
a property of the $\mbox{}^2 g_2$-mode.
Starting from either the left- or right-most column of the time-frequency matrix
we identify and trace the sequence of amplitude peaks within a certain frequency
band given the monotonicity constraint. Appendix \ref{app:gmode} is providing more
details on the reconstruction of the $g$ mode ridge.

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{plots/spectrogram}
 \caption{Spectrogram of the gravitational wave signal {\texttt s20S} sampled at \unit[4096]{Hz}.
   The spectrogram is obtained using data streach of 200 samples overlapping at 90\%
   with each other.} \label{fig:spectrogram}
\end{figure}

We collect the instantaneous frequency $f(t_i)$ corresponding to the ridge $m(t_i)$ for
the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
for values in between the $t_i$. We then use equation (\eqref{model1}) to obtain
estimates of the time evolution of the ratio together with 95\% confidence intervals.
An exemple is given in Figure \ref{fig:ratio} where the red points are the point estimates and
the grey bands represent 95\% confidence bands. Ratio values
computed using the mass and radius values obtained from the simulation code are shown in black.

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/ratio}
 \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ as function of time extracted from the $\mbox{}^2 g_2$-mode of the {\texttt s20S} signal (red points and the 95\% confidence belt in grey) compared to the ratio value derived from the PNS mass and radius given by the simulation code (black points).} \label{fig:ratio}
\end{figure}
